The CrypTool Script: \\ Cryptography, Mathematics and more \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free eLearning program CrypTool\\ (8th edition -- distributed with CrypTool version 1.4.10)} \vspace{60pt}

### 2007-07-16 script/en/chapters/cryptomethods.tex
Update scripts: Corrections according to feedback

% $Id$
% ..............................................................................
%                V E R S C H L U E S S E L U N G S V E R F A H R E N
%
% Schreibregelung / Syntax rules:
%         - Capitel header with all nouns/verbs-words capitalized,
%         - All other headers in the normal lower/upper case manner
%           like sentences without dot at the end.
%
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
%\newpage

@@

\hypertarget{Kapitel_1}{}
\section{Encryption Procedures}
(Bernhard Esslinger, May 1999; Updates Dec. 2001, Feb. 2003, June 2005)

@@

% --------------------------------------------------------------------------
% \newpage
\subsection[Hybrid procedures]
{Hybrid procedures\footnotemark}
\footnotetext{%
Within CrypTool\index{CrypTool} you can get a visualization of this
technique using the menu path {\bf Crypt \textbackslash{} Hybrid}: 
this dialogue shows the single steps and its dependencies with concrete
numbers. Here the asymmetric algorithm RSA and the symmetric algorithm AES are used.
}\index{Hybrid procedure}

@@

\bibitem[Schmeh2003]{Schmeh2003}  \index{Schmeh 2003}
        Klaus Schmeh, \\
        {\em Cryptography and Public Key Infrastructures on the Internet},\\ 
        John Wiley \& Sons Ltd., Chichester 2003. \\
        A considerable, up-to-date, easy to read book, which also 
        considers practical problems such as standardisation or
        real existing software.

>>>>

% $Id$
% ..............................................................................
%                V E R S C H L U E S S E L U N G S V E R F A H R E N
%
% Schreibregelung / Syntax rules English:
%         - Capitel header with all nouns/verbs-words capitalized,
%         - All other headers in the normal lower/upper case manner
%           like sentences, but without dot at the end.
%
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

@@

\hypertarget{Kapitel_1}{}
\section{Encryption Procedures}
(Bernhard Esslinger, May 1999; Updates Dec. 2001, Feb. 2003, June 2005, July 2007)

@@

% --------------------------------------------------------------------------
% \newpage
\subsection[Hybrid procedures]
{Hybrid procedures\footnotemark}
\footnotetext{%
Within CrypTool\index{CrypTool} you can find this technique using the menu
path {\bf Crypt \textbackslash{} Hybrid}: 
There you can follow the single steps and its dependencies with concrete
numbers. The variant with RSA as the asymmetric algorithm is graphically
visualized; the variant with ECC uses the standard dialogs. In both cases
AES is used as the symmetric algorithm.
}\index{Hybrid procedure}

@@

\bibitem[Schmeh2003]{Schmeh2003}  \index{Schmeh 2003}
        Klaus Schmeh, \\
        {\em Cryptography and Public Key Infrastructures on the Internet},\\ 
        John Wiley \& Sons Ltd., Chichester 2003. \\
        An up-to-date, easy to read book, which also considers practical
        problems such as standardisation or real existing software.

### 2007-07-16 script/en/chapters/numbertheory.tex
Update scripts: Corrections according to feedback

Number theory has always been a source of fascination -- for both amateurs and professional mathematicians. In 
contrast to other areas of mathematics, many of the problems and theorems in number theory can be understood 
by non-experts. On the other hand, mathematicians often take a long time to find solutions to the problems or 
prove the theorems. It is therefore one thing to pose good questions but quite another matter to find the answer. One 
example of this is what is known as Fermat's Last (or large) theorem\footnote{One of the things you learn in mathematics at school is Pythagoras theorem, which states the following for a right-angle 
triangle: $a^2 + b^2 = c^2$, where $a$ and $b$ are the lengths of the sides containing the right angle and $c$ is the 
length of the hypotenuse. Fermat famously proposed that $a^n + b^n \not= c^n$ for whole-number exponents $n 
> 2$. Unfortunately, the letter in which Fermat made the claim did not have enough space for him to prove it. 
The theorem was not proved until over 300 years later \cite[p. 433-551]{Wiles1994}\index{Wiles, Andrew}. }.

@@

\begin{itemize}
\item the addition table\footnote{Comment on subtraction modulo 5: \\
$2 - 4 \equiv -2 \equiv 3{\rm ~mod~}5.$\\
It is therefore not true modulo $5$ that $-2 = 2$ (see also \hyperlink{Appendix_C}{Appendix C of this chapter}). }
${\rm ~(mod~ } 5)$ and
\item the multiplication tables\footnote{\label{ftn-mod6}Comment on division modulo $6$:\index{Division modulo $n$}\\
Due to the special role of zero as the identity for addition, division by zero is not permitted:\\
for all $a$ it is $a*0=0,$ because $a*0=a*(0+0)=a*0+a*0.$ Obviously $0$ has no inverse regarding the
multiplication, because if there would be one, it must be $ 0= 0*0^{-1} =1.$ Also see footnote \ref{ftn-zth-divmodn}.}
for mod $5$ and mod $6$.
\end{itemize}

@@

Modular powers work in the same way as modular addition and modular multiplication:
$$ 3^2 \equiv 9 \equiv 4 {\rm ~(mod~} 5). $$
Even consecutive powers work in the same way:

{\bf Example 1:}
$$ (4^3)^2 \equiv 64^2 \equiv 4096 \equiv 1 {\rm ~(mod~} 5). $$
\begin{quote}
(1) We can speed up\footnote{The time required to calculate the multiplication 
of two numbers normally depends on the length of the numbers.  We can observe
this if we use the school method to calculate, for instance, $474*228$.  The 
time required increases 
%quadratically
in a quadratic square manner , because we need to multiply $3*3$ numbers. The 
numbers become considerably smaller if we reduce the interim result.} the 
calculation by reducing the {\bf interim results} modulo $5$ but we need 
to take care because not everything will then work in the same way as in 
standard arithmetic.
\begin{eqnarray*}
(4^3)^2 & \equiv & (4^3{\rm ~(mod~}5))^2{\rm ~(mod~}5) \nonumber \\
            & \equiv & (64{\rm ~(mod~}5))^2\;{\rm ~(mod~}5) \nonumber \\
            & \equiv & 4^2{\rm ~(mod~}5) \nonumber \\
            & \equiv & 16 \equiv 1 {\rm ~(mod~}5). \nonumber
\end{eqnarray*}

@@

{\bf Example 2:}
$$
3^{28} \equiv 3^{4\:*\:7} \equiv 3^{4\:*\:7{\rm ~(mod~}10)} \equiv 3^8 \equiv 6561 \equiv 5 {\rm ~(mod~}11).
$$

@@

% --------------------------------------------------------------------------
\hypertarget{RSA}{}
\subsubsection{How the RSA procedure works}
\label{RSA}
The individual steps for implementing the \index{RSA} RSA procedure can be described as follows (see \cite[p. 213 ff]{Eckert2003}\index{Eckert 2003} and 
\cite[p. 338 ff]{Sedgewick1990}\index{Sedgewick 1990}). 
Steps 1 to 3 constitute key generation, steps 4 and 5 are the encryption, and steps 
6 and 7 are the decryption:
\begin{itemize}
\item[{\bf 1.}] Select two distinct random prime numbers\footnote{
Compaq introduced the so-called multi-prime method with high marketing 
effort in 2000.
$n$ was the product of two big and one relative small prime: $n=o*p*q.$ 
With theorem \ref{J_of_p1..pk} we get:
$J(n)= (o-1)*(p-1)*(q-1). $ This method did not assert itself yet.\\
One reason probably is, that Compaq claimed a patent \index{Patent} on it. 
Generally there is less understanding in Europe and with the Open Source Initiative
\index{Open Source}, that one can claim patents on algorithms. But there is really
no understanding outside the U.S., that one can get a patent for a special case 
(3 factors) of an algorithm (RSA), although the patent for the general case is
almost expired.}$^,$\footnote{If the two primes $p$ and $q$ are equal then $ (m^{e})^{d} \equiv m $ mod $n$ is not true for all $ m<n$ (although
$ e*d \equiv 1$ mod $J(n)$ is fulfilled). 
Example: If $ n=5^2$ then according to theorem \ref{thm-zth-phinum} it 
is $J(n)=5*4=20, ~ e=3, ~ d=7, ~e*d=21\equiv 1$ mod $J(n).$ 
But it is $ (5^3)^7\equiv 0 $ mod $ 25$. } $p$ and $q$ and calculate 
$n = p*q$\footnote{The GISA\index{GISA} (German Information Security Agency) 
recommends, to choose the prime factors $p$ and $q$ almost the same,
but not too close:
$$ 0.5 < |\log_2 (p) - \log_2 (q) | <30. $$
They recommend to generate the primes independently and check that the 
restriction is fulfilled (see \cite{GISA2002}).
}.\\
The value $n$ is called the RSA modulus\footnote{
In CrypTool\index{CrypTool} the RSA modulo is denoted with a capital ``$N$'' .}.

@@

We will work through the example in 2 variants. The steps 1 to 3 are common for both.\par
{\bf Key generation (steps 1 to 3)}: \label{SrcArith4b} \\
{\bf 1.} $p=503, q=509 \quad (n= 256,027; \; J(n)=(p-1)(q-1)=255,016=2^3*127*251)$\footnote{%
See \hyperlink{AppArith4b}{Appendix E of this chapter} for the source code to factorise
the number $J(n)$ using Mathematica and Pari-GP.
Using CrypTool\index{CrypTool} you can solve this with the 
{\bf Indiv.Procedures \textbackslash{} RSA Cryptosystem \textbackslash{} 
Factorisation of a Number}.}. \\

{\bf 2.} $e=65,537$ (should\footnote{%
                See footnote \ref{foot:Selection-of-e} on page
                \pageref{foot:Selection-of-e}.}
                lie between $509$ and $255,015$ and must\footnote{%
       $e$ cannot, therefore, be $2, 127$ or $251$
       ($65,537 = 2^{16}+1$) ($255,016 = 2^{3}*127*251$).\\
       In real life, $J(n)$ is not factorised but rather the Euclidean
       algorithm is used for the selected e to guarantee that
       ${\rm gcd}(e,J(n))=1$.
                 } be relatively prime to $255,016$).\\

{\bf 3.} $d=231,953$ \\\strut\quad\ (since $e \equiv d^{-1}{\rm ~mod~}J(n): ~65,537*231,953 \equiv 15,201,503,761 \equiv 1
{\rm ~(mod~}255,016)$)\footnote{Other possible combinations of $(e,d)$ include: $(3, 170,011)$, $(5, 204,013)$, $(7, 36,431)$.}.

>>>>

Number theory has always been a source of fascination -- for both amateurs and
professional mathematicians. In contrast to other areas of mathematics, many
of the problems and theorems in number theory can be understood by non-experts.
On the other hand, the solutions to these problems or the prove to the theorems
often resisted to the mathematicians for a very long time.
It is therefore one thing to pose good questions but quite another matter to
find the answer. One example of this is what is known as Fermat's Last (or
large) theorem\footnote{One of the things we learn in mathematics at school
   is Pythagoras' theorem, which states the following for a right-angle
   triangle: $a^2 + b^2 = c^2$, where $a$ and $b$ are the lengths of the sides
   containing the right angle and $c$ is the length of the hypotenuse.
   Fermat famously proposed that $a^n + b^n \not= c^n$
   for $a,b,c \in \mathbb{N}$ and whole-number exponents $n > 2$.
   Unfortunately, the letter in which Fermat made the claim did not have
   enough space for him to prove it. The theorem was not proven until over
   300 years later \cite[p. 433-551]{Wiles1994}\index{Wiles, Andrew}. }.

@@

\begin{itemize}
\item the addition table\footnote{Comment on subtraction modulo 5: \\
$2 - 4 = -2 \equiv 3{\rm ~mod~}5.$\\
It is therefore not true modulo $5$ that $-2 = 2$ (see also \hyperlink{Appendix_C}{Appendix C of this chapter}). }
${\rm ~(mod~ } 5)$ and
\item the multiplication tables\footnote{\label{ftn-mod6}Comment on division modulo $6$:\index{Division modulo $n$}\\
Due to the special role of zero as the identity for addition, division by zero is not permitted:\\
for all $a$ it is $a*0=0,$ because $a*0=a*(0+0)=a*0+a*0.$ Obviously $0$ has no inverse regarding the
multiplication, because if there would be one, it must be $ 0= 0*0^{-1} =1.$ Also see footnote \ref{ftn-zth-divmodn}.}
for mod $5$ and mod $6$.
\end{itemize}

@@

Modular powers work in the same way as modular addition and modular multiplication:
$$ 3^2 = 9 \equiv 4 {\rm ~(mod~} 5). $$
Even consecutive powers work in the same way:

{\bf Example 1:}
$$ (4^3)^2 = 64^2 \equiv 4096 \equiv 1 {\rm ~(mod~} 5). $$
\begin{quote}
(1) We can speed up\footnote{The time required to calculate the multiplication 
of two numbers normally depends on the length of the numbers.  We can observe
this if we use the school method to calculate, for instance, $474*228$.  The 
time required increases 
%quadratically
in a quadratic square manner , because we need to multiply $3*3$ numbers. The 
numbers become considerably smaller if we reduce the interim result.} the 
calculation by reducing the {\bf interim results} modulo $5$ but we need 
to take care because not everything will then work in the same way as in 
standard arithmetic.
\begin{eqnarray*}
(4^3)^2 & \equiv & (4^3{\rm ~(mod~}5))^2{\rm ~(mod~}5) \nonumber \\
            & \equiv & (64{\rm ~(mod~}5))^2\;{\rm ~(mod~}5) \nonumber \\
            & \equiv & 4^2{\rm ~(mod~}5) \nonumber \\
            & \equiv & 16 \equiv 1 {\rm ~(mod~}5). \nonumber
\end{eqnarray*}

@@

{\bf Example 2:}
$$
3^{28} = 3^{4\:*\:7} \equiv 3^{4\:*\:7{\rm ~(mod~}10)} \equiv 3^8 \equiv 6561 \equiv 5 {\rm ~(mod~}11).
$$

@@

% --------------------------------------------------------------------------
\hypertarget{RSA}{}
\subsubsection{How the RSA procedure works}
\label{RSA}
The individual steps for implementing the \index{RSA} RSA procedure can be described as follows (see \cite[p. 213 ff]{Eckert2003}\index{Eckert 2003} and 
\cite[p. 338 ff]{Sedgewick1990}\index{Sedgewick 1990}). 
Steps 1 to 3 constitute key generation, steps 4 and 5 are the encryption, and steps 
6 and 7 are the decryption:
\begin{itemize}
\item[{\bf 1.}] Select two distinct random prime numbers\footnote{
Compaq introduced the so-called multi-prime method with high marketing 
effort in 2000.
$n$ was the product of two big and one relative small prime: $n=o*p*q.$ 
With theorem \ref{J_of_p1..pk} we get:
$J(n)= (o-1)*(p-1)*(q-1). $ This method did not assert itself yet.\\
One reason probably is, that Compaq claimed a patent \index{Patent} on it. 
Generally there is less understanding in Europe and with the Open Source
Initiative\index{Open Source}, that one can claim patents on algorithms.
But there is really no understanding outside the U.S., that one can get
a patent for a special case (3 factors) of an algorithm (RSA), although
the patent for the general case is almost expired.}$^,$\footnote{If the
two primes $p$ and $q$ are equal then $ (m^{e})^{d} \equiv m $ mod $n$ is
not true for all $ m<n$ (although $ e*d \equiv 1$ mod $J(n)$ is fulfilled). 
Example: If $ n=5^2$ then according to theorem \ref{thm-zth-phinum} it 
is $J(n)=5*4=20, ~ e=3, ~ d=7, ~e*d=21\equiv 1$ mod $J(n).$ 
But it is $ (5^3)^7\equiv 0 $ mod $ 25$. } $p$ and $q$ and calculate 
$n = p*q$\footnote{The GISA\index{GISA} (German Information Security Agency) 
recommends, to choose the prime factors $p$ and $q$ almost the same,
but not too close:
$$ 0.5 < |\log_2 (p) - \log_2 (q) | <30. $$
They recommend to generate the primes independently and check that the 
restriction is fulfilled (see \cite{GISA2002}).
}.\\
The value $n$ is called the RSA modulus\footnote{
In CrypTool\index{CrypTool} the RSA modulo is denoted with a capital ``$N$'' .}.

@@

We will work through the example in 2 variants. The steps 1 to 3 are common for both.\par
{\bf Key generation (steps 1 to 3)}: \label{SrcArith4b} \\
{\bf 1.} $p=503,~q=509 \quad (n= 256,027; \; J(n)=(p-1)(q-1)=255,016=2^3*127*251)$\footnote{%
See \hyperlink{AppArith4b}{Appendix E of this chapter} for the source code to factorise
the number $J(n)$ using Mathematica and Pari-GP.
Using CrypTool\index{CrypTool} you can solve this with the 
{\bf Indiv.Procedures \textbackslash{} RSA Cryptosystem \textbackslash{} 
Factorisation of a Number}.}. \\
{\bf 2.} $e=65,537$ \\
\strut\quad\ (should\footnote{%
                See footnote \ref{foot:Selection-of-e} on page
                \pageref{foot:Selection-of-e}.}
                lie between $509$ and $255,015$ and must\footnote{%
       $e$ cannot, therefore, be $2, 127$ or $251$
       ($65,537 = 2^{16}+1$) ($255,016 = 2^{3}*127*251$).\\
       In real life, $J(n)$ is not factorised but rather the Euclidean
       algorithm is used for the selected e to guarantee that
       ${\rm gcd}(e,J(n))=1$.
                 } be relatively prime to $255,016$).\\
{\bf 3.} $d=231,953$ \\
\strut\quad\ (since $e \equiv d^{-1}{\rm ~mod~}J(n): ~65,537*231,953 \equiv 15,201,503,761 \equiv 1
{\rm ~(mod~}255,016)$)\footnote{Other possible combinations of $(e,d)$ include: $(3, 170,011)$, $(5, 204,013)$, $(7, 36,431)$.}.

### 2007-07-16 script/en/chapters/primes.tex
Update scripts: Corrections according to feedback

\newpage
\hypertarget{Kapitel_2}{}
\section{Prime Numbers}
\label{Label_Kapitel_2}
(Bernhard Esslinger, May 1999; Updates Nov. 2000, Dec. 2001, June 2003, May 2005, March 2006, May 2007)

@@

As Euclid discovered, there is an infinite number of primes. However, some
infinite sets are {\em denser} \index{Prime number!density} than others. Within
the set of natural numbers, there is an infinite number of even, uneven and
square numbers.

The following proves that there are more even numbers than square ones:
\begin{itemize}
  \item the size of the $n$th element: \\
    The $n$th element of the even numbers is $2n$; the $n$th element of the
square numbers is $n^2$. Because for all $n>2$: $2n < n^2$, the $n$th even
number occurs much earlier than the $n$th square number.
    Thus the even numbers are distributed more densely and we can say that there
are more even numbers than square ones.
  \item the number of values that are less than or equal to a certain {\em
maximum value} $x$ in ${\mathbb R}$ is: \\
    There are $[x/2]$ such even numbers and $[\sqrt{x}]$ square numbers. Because
for large $x$ the value $x/2$ is much greater than the square root of $2$, we
can again say that there are more even numbers.
\end{itemize}

@@

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|l|}\hline
Dual system   &   Decimal system \\
\hline \hline
$2^{10}$ 	& $1024$ \\
$2^{40}$ 	&  $1.09951\cdot 10^{12}$ \\
$2^{56}$ 	&  $7.20576\cdot 10^{16}$ \\
$2^{64}$ 	&  $1.84467\cdot 10^{19}$ \\
$2^{80}$ 	&  $1.20893\cdot 10^{24}$ \\
$2^{90}$ 	&  $1.23794\cdot 10^{27}$ \\
$2^{112}$ 	&  $5.19230\cdot 10^{33}$ \\
$2^{128}$ 	&  $3.40282\cdot 10^{38}$ \\
$2^{150}$ 	&  $1.42725\cdot 10^{45}$ \\
$2^{160}$ 	&  $1.46150\cdot 10^{48}$ \\
$2^{250}$ 	&  $1.80925\cdot 10^{75}$ \\
$2^{256}$ 	&  $1.15792\cdot 10^{77}$ \\
$2^{320}$ 	&  $2.13599\cdot 10^{96}$ \\
$2^{512}$ 	&  $1.34078\cdot 10^{154}$ \\
$2^{768}$ 	&  $1.55252\cdot 10^{231}$ \\
$2^{1024}$	&  $1.79769\cdot 10^{308}$ \\
$2^{2048}$	&  $3.23170\cdot 10^{616}$ \\ \hline
\end{tabular}
\caption{Special values of the binary and decimal system}
\end{center}
\end{table}

Calculation using GMP, for example:
{\href{http://www.gnu.ai.mit.edu}{\tt http://www.gnu.ai.mit.edu}}.

>>>>

\newpage
\hypertarget{Kapitel_2}{}
\section{Prime Numbers}
\label{Label_Kapitel_2}
(Bernhard Esslinger, May 1999; Updates Nov. 2000, Dec. 2001, June 2003, May 2005, March 2006, June 2007)

@@

As Euclid discovered, there is an infinite number of primes. However,
some infinite sets are {\em denser} \index{Prime number!density} than
others.

Within the set of natural numbers, there is an infinite number of even,
uneven and square numbers. How  to compare the ``density'' of two
infinite sets is shown with even and square numbers.

The following proves that the even numbers are distributed more densely
than square ones:\footnotemark
   \footnotetext{%
   Whereas in colloquial language you often can hear, that ``there are
   more'' even numbers than square ones, mathematicians say, that from
   both there are inifinitely many, that their sets are equvalent to
   $ \mathbb{N}$ (so both are inifinite and countable, i.e. one can
   assign to each even number and to each square number an integer),
   but that the set of even numbers is denser than the set of square
   numbers.
   }
\begin{itemize}
  \item the size of the $n$th element: \\
    The $n$th element of the even numbers is $2n$;
    the $n$th element of the square numbers is $n^2$.
    Because for all $n>2$: $2n < n^2$, the $n$th even number occurs much
    earlier than the $n$th square number.
  \item the numbers of values that are less than or equal to a certain
    {\em maximum value} $x$ in ${\mathbb R}$ are: \\
    There are $\lfloor x/2 \rfloor$ such even numbers and
    $\lfloor \sqrt{x} \rfloor$ square numbers.
    Because for all $x>6$ the value $\lfloor x/2 \rfloor$ is greater
    than the largest integer smaller or equal to the square root of $x$,
    the even numbers are distributed more densely.
\end{itemize}

@@

These values can be used, to evaluate from a key length in bit the
according number of possible keys and the search effort (if assumed, that
e.g. one million keys can be tested within one second).

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|l|}\hline
Dual system   &   Decimal system \\
\hline \hline
$2^{10}$ 	& $1024$ \\
$2^{40}$ 	&  $1.09951\cdot 10^{12}$ \\
$2^{56}$ 	&  $7.20576\cdot 10^{16}$ \\
$2^{64}$ 	&  $1.84467\cdot 10^{19}$ \\
$2^{80}$ 	&  $1.20893\cdot 10^{24}$ \\
$2^{90}$ 	&  $1.23794\cdot 10^{27}$ \\
$2^{112}$ 	&  $5.19230\cdot 10^{33}$ \\
$2^{128}$ 	&  $3.40282\cdot 10^{38}$ \\
$2^{150}$ 	&  $1.42725\cdot 10^{45}$ \\
$2^{160}$ 	&  $1.46150\cdot 10^{48}$ \\
$2^{192}$ 	&   $6,27710\cdot 10^{57}$ \\
$2^{250}$ 	&  $1.80925\cdot 10^{75}$ \\
$2^{256}$ 	&  $1.15792\cdot 10^{77}$ \\
$2^{320}$ 	&  $2.13599\cdot 10^{96}$ \\
$2^{512}$ 	&  $1.34078\cdot 10^{154}$ \\
$2^{768}$ 	&  $1.55252\cdot 10^{231}$ \\
$2^{1024}$	&  $1.79769\cdot 10^{308}$ \\
$2^{2048}$	&  $3.23170\cdot 10^{616}$ \\ \hline
\end{tabular}
\caption{Special values of the binary and decimal system}
\end{center}
\end{table}

Such tables can easily be calculated with computer algebra systems.
Here a code sample for the CAS system Pari-GP\index{Pari-GP}:
\vspace{-2ex}
\begin{verbatim}
v2=[56, 64, 80, 90, 112, 128, 192, 256];
default(format,"e0.6");
for ( i=1,8, o=v2[i]; print(o, " --- ", 1.*2^o) );
\end{verbatim}
